---
title: "ehabwasel_3"
author: "Ehab"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Fundamental Techniques in Data Science with R - Practical 3
## Introduction
In this practical you will learn about exploratory data analysis (EDA). EDA involves inspecting, transforming and visualising data with the goal of finding interesting features to substantiate one or more research questions. This practical will draw on skills you learned in week 1 and week 2, while introducing you to some new material like visualisation with ggplot().

You will need the following packages for this tutorial.


```{r}
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(weathermetrics)
library(ggpubr)


```
```{r}
str(mpg)
```

# Import dataset gpm from ggplot2 package
```{r}
head(mpg)
```
```{r}
glimpse(mpg)
```
## Exploratory data analysis

```{r}
mpg %>% 
  group_by(class) %>% 
  summarize(
    min = min(hwy),
    mean = mean(hwy),
    sd = sd(hwy),
    max = max(hwy),
    n = n()
  ) %>%
  kbl()  %>% 
  kable_styling(latex_options = c ("striped","hover"), full_width = F)
```


```{r}
mpg  <- mpg %>% 
  mutate_if(is.character , as.factor)

glimpse(mpg)
```


```{r}
mpg %>% 
  group_by(year) %>%
  summarise( 
            min = min(displ),
             mean = mean(displ), 
             median  = median(displ),
             q1 = quantile(displ,0.25),
             q3 = quantile(displ , 0.75),
             max = max(displ) ,
             interqurtile = q3 - q1)%>%
kbl() %>% 
  kable_styling(latex_options = c ("striped", "hover") , full_width = F)

```


```{r}
hist(mpg$displ)
```


```{r}
barplot(table(mpg$cyl))
```


```{r}
plot(x = mpg$displ, y = mpg$hwy,
     xlab = "Highway mpg",
     ylab = "Engine displacement (L)")
```
```{r}
ggplot(iris) + # Data
  geom_point(mapping = aes(x = Sepal.Length, # Variable on the x-axis
                           y = Sepal.Width, # Variable on the y-axis
                           colour = Species)) + # Legend 
  labs(x = "Sepal Length",
       y = "Sepal Width", 
       title = "Relationship between Sepal Length and Width by Species") +
  coord_cartesian() + # Default standard for mapping x and y
  facet_wrap(~Species) + # Splits plot by the species variable
  theme_bw() # Sets the background theme
```
## Visualising distributions of single variables
```{r}

ggplot(mpg) +
  geom_bar(aes(x = drv))
```


```{r}
ggplot(mpg) +              # list the dataset
  geom_bar(aes(x = class)) # geom_bar for a barplot, and the 'class' variable on the x-axis. 
```
```{r}
ggplot(mpg) +
  geom_bar(aes(x = class)) +
  theme_bw() # Here we add the theme
```
## Continuous variables
Histograms are often used to visualise the distribution of a continuous variable. Below is an example of how to visualise the distribution of cty - the number of city miles per gallon in the mpg dataset.

geom_hist transforms the x-axis into equally spaced “bins” and the height of each bar on the y-axis is the number of observations falling in each bin.



```{r}

```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, 
                 y = hwy))
```


```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, 
                 y = hwy, 
                 colour = class))
```


```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, 
                 y = hwy, 
                 colour = trans)) +
  labs(x = "Engine Displacement (L)",
       y = "Highway miles per gallon",) +
  theme_bw()
```


```{r}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Engine Displacement (L)",
       y = "Highway miles per gallon",) +
  theme_bw()
```


```{r}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                colour = class)) +
  geom_point() +
  geom_smooth(method = lm, 
              se = FALSE) +
  labs(x = "Engine Displacement (L)",
       y = "Highway miles per gallon",) +
  theme_bw()
```


```{r}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                colour = class)) +
  geom_point() +
  geom_smooth(method = lm, 
              se = FALSE) +
  labs(x = "Engine Displacement (L)",
       y = "Highway miles per gallon",) +
  theme_bw()
```


```{r}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                colour = class)) +
  geom_point() +
  geom_smooth(method = lm, 
              se = FALSE) +
  labs(x = "Engine Displacement (L)",
       y = "Highway miles per gallon",) +
  theme_bw()
```


```{r}
ggplot(mpg) +
  geom_boxplot(aes(x = drv, y = hwy)) +
  coord_flip()
```


```{r}
mpg <- mpg %>% 
  mutate(trans = fct_collapse(trans,
                             "auto" = c("auto(av)", 
                                        "auto(l3)", 
                                        "auto(l4)", 
                                        "auto(l5)", 
                                        "auto(l6)", 
                                        "auto(s4)", 
                                        "auto(s5)", 
                                        "auto(s6)"),
                             "manual" = c("manual(m5)", 
                                          "manual(m6)"))) 
```


```{r}
ggplot(mpg) +
  geom_boxplot(aes(x = drv, 
                   y = cty, 
                   colour = trans))
```


```{r}
mpg %>% 
    mutate(trans = fct_collapse(trans,
                             "auto" = c("auto(av)", 
                                        "auto(l3)", 
                                        "auto(l4)", 
                                        "auto(l5)", 
                                        "auto(l6)", 
                                        "auto(s4)", 
                                        "auto(s5)", 
                                        "auto(s6)"),
                             "manual" = c("manual(m5)", 
                                          "manual(m6)"))) %>% 
  ggplot() +
  geom_boxplot(aes(x = drv, 
                   y = cty, 
                   colour = trans))
```


```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, 
                 y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```


```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, 
                 y = hwy)) + 
  facet_grid(drv ~ cyl)
```


```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, 
                 y = hwy)) + 
  facet_wrap(~manufacturer, nrow = 3)
```


```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, 
                 y = hwy)) + 
  labs(x = "Engine displacement (litres)", 
       y = "Highway miles per gallon") +
  facet_wrap(~manufacturer, nrow = 3) 
```
## Arranging multiple plots

```{r}
# plot A
p1 <- ggplot(data = mpg, aes(class, fill = class)) +
        geom_bar() +
        labs(x = "Vehicle class",
              y = "Count") +
        coord_flip() +  
        theme_bw()

# plot B
p2 <- ggplot(data = mpg) + 
        geom_boxplot(mapping = aes(x = class, y = hwy, colour = class)) +
        labs(x = "Vehicle class",
             y = "Highway mpg") +
        coord_flip() +
        theme_bw()
# plot C
p3 <- ggplot(data = mpg) + 
        geom_jitter(mapping = aes(x = cty, y = hwy, colour = class)) +
        labs(x = "Vehicle class",
             y = "Highway mpg") +
        theme_bw()
```


```{r}
ggarrange(p1, p2, p3, 
          ncol = 2, 
          nrow = 2)
```
```{r}
ggarrange(p1, p2, p3, 
          nrow = 2, 
          ncol = 2, 
          vjust = 0.5, 
          common.legend = TRUE,
          legend = "bottom")

```

