---
title: "Practical 1: Introduction"
author: "ehab_wasel"
date: "`r Sys.Date()`"
output: html_document

---
## step 1 : Load the required packages:
```{r setup, include=FALSE}
library(dplyr)
library(readr)
library(knitr)
```
## Step 2: Basic Mathematical Operations
```{r code_exercises}
# 1. Create an object 'a' with value 1
a <- 1 

# 2. Verify that 'a' is stored as 1
a # Prints the value of 'a'

# 3. Square 'a'
a_squared <- a^2 

# 4. Assign a + a to 'b' and check if 'b' is equal to a + a
b <- a + a
b == a + a # Should return TRUE

# 5. Square 'b'
b_squared <- b^2 

# 6. Multiply the result of question 5 by a over b
c <- (b_squared * a) / b 

# 7. Assign the result to 'c'
c

# 8. Take the square root of 'c' to the power 'b'
sqrt_c_b <- sqrt(c^b) 

# 9. Multiply the result by a over (b to the power of 6)
result <- (sqrt_c_b * a) / (b^6)

# 10. Round the result to 3 decimal places
rounded_result <- round(result, 3)
rounded_result # Print final result
```
## Step 3: Using Built-in Functions
```{r functions}
# Create a sequence from 12 to 24
seq_12_24 <- seq(12, 24)
seq_12_24 

# Sum the numbers from 28 to 63
sum_28_63 <- sum(28:63)
sum_28_63 

# Find the mean of numbers from 25 to 82
mean_25_82 <- mean(25:82)
mean_25_82 
```{r loading_data ,Step 5: Summarizing and Mutating Data}

# Load the flight data
flight_data <- read_csv("../dataset/flightdata.csv")

# Display the data
flight_data
```{r data_summary}
# Summarize the data
summary(flight_data)
```{r mutate_data}
# Add a new column 'speed'
flight_data2 <- flight_data %>%
  mutate(speed = distance / air_time * 60)

# Display the updated dataset
head(flight_data2)
```{r select_data}
# Select specific columns
flight_data3 <- flight_data2 %>%
  select(year, month, day, speed)

# Display the new data
head(flight_data3)

```{r for_loop}
# Loop to calculate third power divided by 13
for (i in 1:12) {
  result <- (i^3) / 13
  print(result)
}

```{r apply_statement}
# Create an 8x8 matrix
matrix_data <- matrix(1:64, nrow = 8, ncol = 8)

# Calculate variance of each row
row_variances <- apply(matrix_data, MARGIN = 1, FUN = var)
row_variances
```{r save_data}
write_csv(flight_data3, "ehab_wasel_flight_data3.csv")


