---
title: "Practical 2 Solutions"
author: "Ehab"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Set CRAN mirror
options(repos = c(CRAN = "https://cran.r-project.org"))

# Check if the package is installed, and install if not
if (!require(kableExtra)) install.packages('kableExtra')
library(kableExtra)

gss_cat <- forcats::gss_cat
```
# Exploring the Dataset
# Preview the data

```{r}
head(gss_cat)
str(gss_cat)

```
# Tibbles
```{r}
as_tibble(gss_cat)

```
# pipes
```{r}
gss_cat %>% pull(relig) %>% unique()
```

```{r}
gss_cat %>% 
  filter( relig == "Protestant") %>% 
  group_by(year , relig) %>% 
  summarize( tvhours = mean(tvhours, na.rm = TRUE))
```
#Missing values
##Standard missing values
```{r}
gss_cat[1:10,] %>% 
  kable() %>% 
  kable_styling(latex_options = "striped")

```
# Standard Missing valus in Entire dataset
```{r}
anyNA(gss_cat)
sum(is.na(gss_cat))
gss_cat %>% 
  count(rincome)

```

```{r}
gss_cat %>% count(marital)

```
```{r}
gss_cat %>% 
  count(partyid)
```


```{r}
gss_cat %>% 
  count(relig)
```
 
```{r}
gss_cat %>% 
  count(denom)
```
#Data transformation
##Filtering data


```{r}
gss_cat %>% 
  filter(marital == "Married")
```


```{r}
gss_cat %>% 
  filter(marital == "Divorced"| marital == " Widowed")
```
# Arranging data

```{r}
gss_cat %>%  arrange(tvhours)
```
```{r}
gss_cat %>% arrange(desc(tvhours))

```


```{r}
gss_cat %>% 
  filter(marital == "Married") %>% 
  arrange(desc(tvhours))
  
```


```{r}
gss_cat %>% 
  count(relig) %>% 
  arrange(n)
```
# Summarizing data
## How many hours of tv on average are watched by people of different religions?

```{r}
gss_cat %>%
  group_by(relig) %>% 
  summarise(tvhours = mean(tvhours, na.rm = TRUE))
```

#Strings
```{r}
str_c("x", "y", "z") 

str_c("x", "y", "z", sep = ", ") 
```


```{r}
str_c(c("x", "y", "z"), collapse = "")
```
#Subsetting strings
```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)
```


```{r}
x <- c("apple", "banana", "pear")

str_view(x, "an")
str_view(x, ".a.")
```


```{r}
x <- c("apple", "banana", "pear")

str_view(x, "^a")
str_view(x, "a$")
```


```{r}
x <- c("apple pie", "apple", "apple cake")

str_view(x, "^apple$")
```
#Factors

```{r}
x <- c("Dec", "Apr", "Jam", "Mar")

sort(x)
```


```{r}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
y <- factor(x, levels = month_levels)

print(y)
sort(y)
```
```{r}
gss_cat %>%
  count(race)
```


```{r}

```


```{r}
relig_summary <- gss_cat %>% # Assign the results to object so we can use it later
  group_by(relig) %>% # Group by religion
  summarise(
    age = mean(age, na.rm = TRUE), # Create a mean for age
    tvhours = mean(tvhours, na.rm = TRUE), # Create a mean for tv hours
    n = n())
relig_summary
```


```{r}
relig_summary %>%
  mutate(relig = fct_reorder(relig, tvhours)) %>%
  ggplot(aes(tvhours, relig)) +
    geom_point() +
    xlab("Number of TV hours watched") +
    ylab("Religion") +
    theme_minimal()
```


```{r}
rincome_summary <- gss_cat %>%
  group_by(rincome) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    n = n())

rincome_summary %>% 
  mutate(rincome = fct_reorder(rincome, age)) %>% 
  ggplot(aes(age, rincome)) +
    geom_point() +
    xlab("Average age") +
    ylab("Income") +
    theme_minimal()
```


```{r}
rincome_summary %>% 
  ggplot(aes(age, fct_relevel(rincome, "Not applicable"))) +
  geom_point() +
    xlab("Average age") +
    ylab("Income") +
    theme_minimal()
```
# Since `rincome` already has a principled order (increasing salary), it doesn't make sense to change this. Typically, we only use `fct_reorder()` to rearrange factors that are arbitrarily ordered to something we prefer or that is more intuitive.*
```{r}
gss_cat %>%
  mutate(marital = fct_infreq(marital)) %>% 
  ggplot(aes(marital)) +
    geom_bar(col = "lightblue", fill = "lightblue") +
    theme_minimal()

```


```{r}
gss_cat %>%
  mutate(race = fct_infreq(race)) %>% 
  ggplot(aes(race)) +
    geom_bar(col = "lightblue", fill = "lightblue") +
    theme_minimal()
```


```{r}
gss_cat %>%
  mutate(race = fct_infreq(race)) %>% 
  ggplot(aes(race)) +
    geom_bar(col = "lightblue", fill = "lightblue") +
    theme_minimal()

```

#Modifying factor levels
```{r}
gss_cat %>%
  mutate(partyid = fct_recode(partyid,
                              "Republican" = "Strong republican", 
                              "Republican" = "Not str republican",
                              "Democrat" = "Strong democrat",
                              "Democrat" = "Not str democrat",
                              "Independent" = "Ind,near rep",
                              "Independent" = "Ind,near dem"
                              )) %>% 
  count(partyid)
```


```{r}
gss_cat %>%
  mutate(partyid = fct_recode(partyid,
    NULL = "No answer",
    NULL = "Don't know"
    )) %>%
  count(partyid)
```

```{r}
gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
    NULL = c("No answer", "Don't know"),
    Other = c("Other party"),
    Republican = c("Strong republican", "Not str republican"),
    Independent = c("Ind,near rep", "Independent", "Ind,near dem"),
    Democrat = c("Not str democrat", "Strong democrat")
  )) %>%
  count(partyid)
```


```{r}
gss_cat %>%
  mutate(rincome = fct_collapse(rincome,
    "0 to $10,000" = c("Lt $1000","$1000 to 2999","$3000 to 3999","$4000 to 4999","$5000 to 5999","$6000 to 6999","$7000 to 7999","$8000 to 9999"),
    "$10,000 to $20,000" = c("$10000 - 14999","$15000 - 19999","$20000 - 24999","$25000 or more"),
    NULL = c("Don't know","No answer", "Not applicable","Refused")
    )) %>%  
  count(rincome)
```


```{r}
gss_cat %>%
  mutate(relig = fct_lump(relig)) %>%
  count(relig)
```


```{r}
gss_cat %>%
  mutate(relig = fct_lump(relig, n = 10)) %>%
  count(relig, sort = TRUE)
```
```{r}
gss_cat %>%
  mutate(marital = fct_lump(marital)) %>%
  count(marital, sort = TRUE)
```
#Creating tables

```{r}
gss_sample <- gss_cat[sample(nrow(gss_cat), 10),]
kable(gss_sample)
```


```{r}
select(gss_sample, year, age, race, rincome) %>%
  kable()
```

```{r}
kable(gss_sample) %>% kable_styling(latex_options = "striped")
```


```{r}
kable(gss_sample) %>%
  kable_styling(latex_options = "striped",
                font_size = 8)    
```


```{r}
kable(gss_sample) %>%
  kable_classic %>%                        
  kable_styling(latex_options = "striped",
                font_size =12)
```


```{r}
gss_cat %>% 
  mutate(partyid = fct_recode(partyid,
                              "Republican" = "Strong republican", 
                              "Republican" = "Not str republican",
                              "Democrat" = "Strong democrat",
                              "Democrat" = "Not str democrat",
                              "Independent" = "Ind,near rep",
                              "Independent" = "Ind,near dem"
                              )) %>% 
  group_by(partyid) %>%
  summarise(n=n()) %>% 
  kbl() %>% 
  kable_paper(bootstrap_options = "striped", full_width = F)
```

```{r}
library(rmarkdown)
render("ehabwasel_2.Rmd", output_file = "ehabwasel_2.html")
```



